---
title: "MAPS Majors Historical Data"
author: "Jason Preszler"
date: "September 16, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(purrr)
library(stringr)
library(ggplot2)
library(knitr)
library(kableExtra)

my_kable <- function(df){
  kable(df,"html") %>% 
    kable_styling( bootstrap_options = 
          c("striped", "hover","condensed", "responsive"), full_width=FALSE) 
}

#maj-min-cleaned.csv is from gitlab peak-nav, cleaning code there
majorHist <- read.csv("maj-min-cleaned-2013.csv", header = TRUE) %>% mutate(id = 1:length(Grad.Year))

#reshape along both majors and minors for counting/graphing 
mhLong <- gather(majorHist, key=minorNum, value = minor, -c(id,Grad.Year, Major1,Major2,Major3)) %>% filter(!is.na(minor)) %>% gather(key=majorNum, value=major, -c(id,Grad.Year,minorNum,minor)) %>% filter(!is.na(major))

#majors in maps
mm <- c("MAT","MATPH","MATCS","MAPHD")
mmNames <- c("Mathematics", "Math-Physics", "Math-CS", "Math-Physics ???")

mmDF <- data.frame(code = mm, name=mmNames) %>% arrange(mm)
```

The purpose of this document is to investigate historical trends of majors within the MAPS department and see if there are relationships with certain majors. We focus on *Graduation Years* 2011-2017, and it should be noted that the graduation year is often one year earlier than their "class year" since, for example, a graduation year of 2017 includes students receiving degrees in Summer or Fall of 2017 and Spring of 2018. 

The following table gives the major codes (used by the registrar and institutional research) an names for each major. Throughout this document we will primarily use the codes to refer to each major.

```{r mmTab, echo=FALSE}
my_kable(mmDF)
```

The current catalog (2018-2019) lists 10 NSM majors. Other peaks have similar numbers of majors - 12 in HFA, 7 in SSH, and 6 in PSE. 

## Data Acquisition and Preparation
The data was provided by the Registrar and Director of Institutional Research. It includes all students with graduation year 2013 or later. The original dataset included some students from earlier, pre-PEAK, years who have been removed. Identifying information including name and ID number have been removed, as have additional variables of graduation date, graduation term, degree (BA or BS), and "Concentration". A generic ID was added to to connect the multiple minors and majors together (as a single student) when reshaping data but this has no connection to the student ID. Leading and trailing white-space from all non-numeric variables was also removed which was necessary since some minor codes has leading spaces during earlier years.

## Overview of Data
The data shows that the `r length(majorHist$id)` students received `r length(unique(mhLong$minor))` different minors and `r length(unique(mhLong$major))` different majors. 

## Major Counts over Time
First we look at the overall count of students received any MAPS major, then we see how those counts are spread over time.

```{r majorCnts, echo=FALSE}
filter(mhLong, major %in% mm) %>% group_by(major) %>% summarise(count = n()) %>% my_kable()

xtabs(~major+Grad.Year, data=filter(mhLong, major %in% mm)) %>% my_kable() 
```


```{r minorGraph, echo=FALSE}
filter(mhLong, major %in% mm) %>% group_by(Grad.Year,major) %>% summarise(count = n()) %>% ggplot(aes(x=Grad.Year, y=count, col=major)) + scale_color_manual(values = c("#e41a1c","#377eb8","#4daf4a","#984ea3","#ff7f00","#ffff33","#a65628","#f781bf"))+ geom_line()+theme_bw()+ggtitle("MAPS Major Counts, Graduation Years 2013-2017")
```

## Major-Minor Associations
Below we have a table of the majors in MAPS and minors with more than 4 students earning that major/minor combination. 

```{r majorAssoc, echo=FALSE}
mhlW <- mhLong %>% mutate(majorNum = ifelse(major %in% mm, "Major1", "AddMajor")) %>% filter(majorNum=="Major1") 
mhlW %>% group_by(major, minor) %>% summarise(count=n()) %>% arrange(desc(count)) %>% filter(count>4) %>% my_kable() 
```

## Chord Diagram of MAPS Majors
```{r chords, echo=FALSE, message=FALSE, warning=FALSE}
library(RColorBrewer)
library(chorddiag)

majMin <- select(majorHist, Major1, Major2, Major3, Minor1, Minor2, Minor3, Grad.Year) %>%
  gather(key=MajNum, value = MAJOR, -c(Minor1,Minor2, Minor3, Grad.Year)) %>% select(-MajNum) %>%
  gather(key=MinorNum, value=minor, -c(MAJOR, Grad.Year)) %>%
  select(-MinorNum) %>%
  mutate(MAJOR = str_trim(MAJOR, side="both"), minor = str_to_lower(str_trim(minor, side="both")))

majMin <- filter(majMin, !is.na(minor), minor!="", !is.na(MAJOR), MAJOR %in% mm)

qual_cols = brewer.pal.info[brewer.pal.info$category == 'qual',]
many_colors = unlist(mapply(brewer.pal, qual_cols$maxcolors, rownames(qual_cols)))

chorddiag(
  xtabs(~MAJOR+minor, data = majMin,
        drop.unused.levels = TRUE),
  showTicks = FALSE, groupColors = many_colors, type = "bipartite"
)
```