---
title: "Course Distribution"
author: "Jason Preszler"
date: "May 4, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(stringr)
library(knitr)
```

The current (2017-2018) undergraduate catalog was used to build a list of subjects, course numbers, names, and urls to course descriptions. Here's what the dataset looks like:
```{r loadClass}
classes <- read.csv("data/class-list-2017-2018.csv", header=TRUE)
kable(head(classes))
```

Note that there are `r length(classes$sub)` courses in the catalog among the `r length(unique(classes$sub))` subjects.

Below is a table of courses by subject code (note this is not aligned with a department) and number of courses by subject and course level.

## Courses by Subject
```{r subNum}
classes %>% group_by(sub) %>% summarise(count=n()) %>% kable()
```

Notice the each of Biology and History acount for almost $10\%$ of the total courses.

## Course by Subject and Level
```{r subLevel}
#add level variable
classes <- classes %>% mutate(level=paste0(str_extract(number, '[:digit:]'),"00"))
group_by(classes, sub, level)%>%summarise(count=n()) %>% kable()
```

